Procedural and noisy meshes can have usable UV atlases while still producing visible physically based rendering (PBR) baking artifacts: residual hot spots, mip leakage, cross-channel seams, and unstable relighting. We introduce a differentiable PBR baking-error atlas that treats held-out relighting error as the deployment signal for atlas repair. The method bakes albedo, normal, roughness, and metallic channels through a differentiable PBR renderer, attributes residuals back to faces and charts, repairs only high-residual local charts, reallocates a fixed texel budget with residual-aware demand, and deploys the candidate atlas only if a matched-control C5 gate accepts it. Across 20 headline asset/baseline combinations (12 main and 8 transfer), the gate triggers in three cases: spot/PartUV, bunny/PartUV, and a warped-cylinder procedural transfer case; a supplemental procedural sweep adds one additional non-overlapping acceptance (\cref{tab:transfer_extended}, proc\_crumpled\_box; warped cylinder is shared with the headline transfer set and not double-counted). In those accepted cases, relit peak signal-to-noise ratio improves by +14.67 dB, +10.23 dB, and +10.74 dB, respectively; the remaining 17 rows are rolled back to the baseline atlas. A multi-seed proposal/gate/test split rerun on the accepted PartUV cases shows the gains hold on the held-out test split (spot $+12.58 \pm 0.83$ dB, bunny $+11.84 \pm 0.89$ dB; rollback controls remain at $\Delta = 0.00 \pm 0.00$ across three seeds). An expanded 18-mesh real-mesh sweep covering Fantasia3D, threestudio, threestudio-Trellis, and Objaverse init shapes yields 5 PartUV-initialized confirmed accepts (3 multi-seed + 2 single-seed), one seed-sensitive borderline that is not counted in the headline, and 12 rollbacks (28\% headline acceptance rate; 33\% including the borderline). An independent 12-asset Polyhaven CC0 external-validity track that replaces the synthetic voronoi oracle with \emph{authored} albedo/normal/metallicRoughness maps yields one further accept (\texttt{Barrel\_01}, +5.30 dB) and 11 correct rollbacks, confirming the residual trigger fires on externally authored PBR and not only on the synthetic oracle. Three of the six new accepts (\texttt{ts\_potion}, \texttt{ts\_nascar}, \texttt{ts\_cabin}) are confirmed across n=3 split seeds with held-out test-split mean$\pm$std PSNR gains of $+7.96 \pm 0.78$, $+2.06 \pm 1.82$, and $+3.72 \pm 0.41$ dB respectively. This is the central robustness claim: PartUV-initialized real meshes more often expose residual-backed atlas leverage than the controlled headline assets, and the new accepts hold under split-seed randomization. The matched-control study shows that the gains are not explained by atlas size, padding, or chart count, but disappear when distortion tail and packing utilization are fully locked. The method is therefore not a universal UV optimizer; it is a selective, rollback-gated way to trade localized UV proxy metrics for substantially better downstream PBR fidelity.
