\section{Limitations and future work}
\label{sec:limitations}

The method is limited to a single UV atlas shared by all PBR channels. This matches the target DCC workflow, but it excludes multi-atlas, UDIM-heavy, or per-channel UV strategies that could improve appearance while changing production constraints. The method also assumes that disjoint proposal, gate, and test relighting sets are available or can be approximated; when the material source is itself unreliable, residual attribution may mix atlas error with material-prediction error.

The generated-asset evaluation remains limited. The eight headline transfer meshes are procedural noisy shapes, and the real-mesh evidence covers five SDS-pipeline init shapes from two independent repositories (Fantasia3D \cite{chen2023fantasia3d} and threestudio \cite{guo2023threestudio}); two accept (\texttt{f3d\_animal}, \texttt{ts\_animal}) and three roll back. This is enough to show the same C1--C5 mechanism transfers to independently generated meshes, but it is not a statistical claim about Trellis-3D \cite{xiang2024_trellis_structured_3d_latents}, GET3D \cite{gao2022_get3d_generative_model}, DreamFusion-style \cite{poole2023_dreamfusion_text_to_3d}, or broader production generated outputs. A larger real generated-mesh sweep, including the Objaverse curated subset and Polyhaven that timed out under our per-mesh PartUV preprocessing budget, remains future work.

Several components need cleaner isolation. C4 seam coupling has engineering value as a diagnostic, but its current A5/A9 ablations show zero isolated signal. The explicit mip term in C3 is also weak in the current table, even though residual-aware allocation helps. Future experiments should use seam-specific stress assets, verified ablation hooks, and controlled mip-frequency cases rather than relying on the main spot/bunny assets alone.

Robustness is also bounded. The light, view, and noise sweeps are demonstrated for both spot and bunny PartUV (\cref{tab:robustness,tab:robust_extended}), while the new leakage-control rerun is intentionally limited to key accept/rollback cases with multiple split seeds. Extending both robustness and multi-seed leakage checks to the full 12-case table, warped cylinder, and additional generated meshes is left for future work, and the noise anomaly at extreme $\sigma$ is intentionally retained rather than smoothed away.

The matched-control results also limit the claim. The method's gain disappears when distortion tail and utilization are fully locked. This is not a hidden bug, but it means the method is unsuitable for pipelines that require strict preservation of those classical UV metrics. Finally, dynamic or 4D assets are not evaluated. Temporal stability of the residual atlas, chart edits, and texel allocation remains open.
